\mytitle

### 2020-05-28 paper.tex
Edited to end of Sect 2.

\section{Introduction}
% Todo
The goal in the wind farm layout optimization (WFLO)
problem~\cite{MOSETTI1994105} is to locate wind turbines within a
predefined area to maximize the total power harvested from the wind
stream.  While physical considerations based on the diameter of the
blades prevent turbines being placed too close to each other, turbine
location decisions influence the total power production due to
interference effects (i.e., wakes) generated by upstream turbines.
These inter-turbine effects are best captured by Jensen's wake model~\cite{jensen1983note} when superimposed
using a sum-of-squares expression~\cite{Zhang2014}. However,
this wake modeling yields intractable optimization problems
that must be approximated using either quadratic or linear functions~\cite{turner2014new}.
An alternative formulation of WFLO relies on a less accurate wake model, which uses linear superpositioning
to capture multiple wakes~\cite{donovan2005wind}.
The latter can be captured using quadratic
constraints and objective functions. However, exact optimization
approaches that have been proposed for the alternative WFLO formulation~\cite{Zhang2014,kuo2016wind,donovan2005wind,fagerfjall2010optimizing,archer2011wind,sorkhabi2018constrained}
represent the quadratic relations using approximate linear models,
%that approximate wake dependencies,
as solving quadratic programming problems
is computationally challenging compared to linear models.

In this work, we introduce two novel quadratic WFLO models. The first
model represents WFLO as a quadratic constrained optimization
problem (QCOP).  The QCOP can be solved using state-of-the-art mathematical solvers
that have been recently extended for such problems, e.g., Gurobi~\cite{gurobi}.  Our
second model is a quadratic unconstrained binary optimization
(QUBO) representation of the WFLO problem.  Such a formulation enables the use of
nascent specialized optimization hardware tailored to solve QUBOs: we use Fujitsu's Digital Annealer (DA) \cite{aramon2019physics} here.
%, a hardware-based
%QUBO solver, to demonstrate the strength of our approach.

The main
contributions of this paper are:
\begin{enumerate} 
\item We propose a novel quadratic modeling framework for WFLO that encompasses both constrained and unconstrained quadratic optimization models.
 %that captures wake 
%effects directly through their linear superposition. 
%without the need for approximations and ad-hoc solutions,
\item We prove that the QCOP variation of WFLO is $\mathcal{NP}$-hard.
\item Through numerical experiments on two instantiations of our quadratic framework using Fujitsu's Digital Annealer and a commercial software-based quadratic solver (Gurobi), we demonstrate state-of-the-art performance in, respectively, quickly finding high-quality solutions and finding and proving optimal solutions over a longer run-time.
\end{enumerate}

Our experiments are focused on the performance of the two approaches
compared to 
existing optimization methods. To this end,
we experiment with 12 standard WFLO benchmark problems (c.f.~\cite{turner2014new})
and show that solving quadratic WFLO models
using Gurobi and the DA
achieves state-of-the-art performance.
%, often outperforming existing solutions when provided sufficient run-time.
Furthermore, we observe that the DA often provides solutions of
similar quality to software-based approaches but with two orders of
magnitude lower \todo{Eldan: shorter instead of lower?} run-time.
%The ability to quickly obtain competitive
%solutions on large WFLO instances can be employed to solve multiple
%wind farm design problems.

@@

%\paragraph{Wind Farm as a Grid}
We consider a square wind farm that we model as an $n\times n$
grid with $n$ being the number of cells on each axis.  The cells have
an area of $c^2$ (in square meters), i.e., if $c=100$ meters and
$n=20$, the farm area will be $2km\times 2km$.  We
assume that the terrain is flat and that we have no constraints
related to the terrain.  Furthermore, we do not account for the
influence of noise on the wind farm's surroundings. The wind farm
literature treats the more complex farm models that capture non-flat
terrain~\cite{song2015lazy,kuo2016wind}, multi-typed
turbines~\cite{feng2017design}, noise
effects~\cite{Zhang2014,sorkhabi2016impact,yin2014multi}, and complex
objectives (including turbine installation and maintenance
costs)~\cite{lackner2007analytical,sorkhabi2018constrained}.  In this work, we 
choose to focus on a
novel approach for solving WFLO, and hence consider the most
abstract wind farm model. Our model can be extended, in the
future, to include additional problem characteristics as mentioned above.

@@

\paragraph{Proximity Constraints}
Turbines must be placed at least \emph{five rotor diameters}
apart. Depending on the cell size, we shall formulate the
corresponding constraint as part of the optimization model. For
example, if a turbine is placed in the center of a
cell, a cell size of $c = 100$ meters and a rotor diameter
of $40$ meters, requires that two turbines are not placed in adjacent cells. In Figure~\ref{fig:field_model}, the red area
around the turbine represents the forbidden cells.

@@

Let $u_{id}$ be the wind velocity at turbine $i = 1,\ldots, m$ for wind regime $d\in\mathcal{D}$.\footnote{Turbines can adjust their orientation according to the current wind 
regime.}  Given the interference of the wakes, the velocity of turbine $i$ is not equivalent to the free wind speed of $d$.
Rather, $u_{id}$ can be computed using the following equation~\cite{Zhang2014}:
\begin{equation}
u_{id} = u_{id\infty} \Bigg[1 - \sqrt{\sum_{j\in\mathcal{U}_{id}} \bigg( 1-\frac{u_{ijd}}{u_{id\infty}} \bigg)^2}  \Bigg], \label{eq:ss}
\end{equation} with $u_{id\infty}$ being the free wind speed (the second component of regime $d$), $\mathcal{U}_{id}$ corresponding to the set of turbines  that are upstream to $i$ 
under wind regime $d$, and $u_{ijd}$ being the wake-reduced speed at $i$ due to upstream turbine $j$. 
The reduced speed, $u_{ijd}$, is computed using the distance $x$ between the two turbines ($i,j$)  and turbine and terrain specifications that yield the cone opening radius, $r$. The set of upstream turbines, $\mathcal{U}_{id}$ is computed  
by rotating the grid wrt to the current wind regime $d \in \mathcal{D}$ and identifying the turbines, $j$, located upstream of turbine $i$ given $d$. The expression in Eq. (\ref{eq:ss}) is referred to as the \emph{sum-of-squares} (SS) model for 
the total expected power in presence of wakes~\cite{Zhang2014}.  We can now write 
the sum-of-squares (SS) expected power of a wind farm with $m$ turbines:
%~(\cite{Zhang2014}):
\begin{equation}
  E_{SS} = \sum_{i=1}^m \sum_{d\in\mathcal{D}} \frac{1}{3} u_{id}^3p_d.\label{eq:ess}
\end{equation}

@@

To overcome this limitation, Rivas et al.\ \cite{rivas2009solving} used simulated annealing (SA), 
a neighborhood search method that allows visits to infeasible solutions. 
This creates a well-connected solution space that leads the algorithm to escape 
local optima. The SA method was shown to be superior to the approach of Ozturk \& Norman~\cite{ozturk2004heuristic} 
and to the genetic method proposed in Grady et al.\ \cite{grady2005placement}. A drawback of the SA approach 
is its ad-hoc nature. The various components of the algorithm
 must be tailored to the problem at hand and, as a result, small adjustments of the WFLO setting
 (e.g., adding noise considerations), would require major changes to the SA implementation. 
Additional methods and experimental comparisons between various approximate algorithms are available in Samorani~\cite{samorani2013wind}.

>>>>

\section{Introduction}
% Todo
The goal in the wind farm layout optimization (WFLO)
problem~\cite{MOSETTI1994105} is to locate wind turbines within a
predefined area to maximize the total power harvested from the wind
stream.  While physical considerations based on the diameter of the
blades prevent turbines being placed too close to each other, turbine
location decisions influence the total power production due to
interference effects (i.e., wakes) generated by upstream turbines.
These inter-turbine effects are best captured by Jensen's wake model~\cite{jensen1983note} when superimposed
using a sum-of-squares expression~\cite{Zhang2014}. However,
this wake modeling yields intractable optimization problems
that must be approximated using either quadratic or linear functions~\cite{turner2014new}.
An alternative formulation of WFLO relies on a less accurate wake model, using linear superpositioning
to capture multiple wakes~\cite{donovan2005wind}.
The latter can be captured using quadratic
constraints and objective functions. However, exact optimization
approaches that have been proposed for the alternative WFLO formulation~\cite{Zhang2014,kuo2016wind,donovan2005wind,fagerfjall2010optimizing,archer2011wind,sorkhabi2018constrained}
represent the quadratic relations using approximate linear models,
%that approximate wake dependencies,
as solving quadratic programming problems
is computationally challenging compared to linear models.

In this work, we introduce two novel quadratic WFLO models. The first
model represents WFLO as a quadratic constrained optimization
problem (QCOP).  The QCOP can be solved using state-of-the-art mathematical solvers
that have been recently extended for such problems, e.g., Gurobi~\cite{gurobi}.  Our
second model is a quadratic unconstrained binary optimization
(QUBO) representation of the WFLO problem.  Such a formulation enables the use of
nascent specialized optimization hardware tailored to solve QUBOs: we use Fujitsu's Digital Annealer (DA) \cite{aramon2019physics}.
%, a hardware-based
%QUBO solver, to demonstrate the strength of our approach.

The main
contributions of this paper are as follows:
\begin{enumerate} 
\item We propose a novel quadratic modeling framework for WFLO that encompasses both constrained and unconstrained quadratic optimization models.
 %that captures wake 
%effects directly through their linear superposition. 
%without the need for approximations and ad-hoc solutions,
\item We prove that the QCOP variation of WFLO is $\mathcal{NP}$-hard.
\item Through numerical experiments on two instantiations of our quadratic framework using Fujitsu's Digital Annealer and a commercial software-based quadratic solver (Gurobi), we demonstrate state-of-the-art performance in, respectively, quickly finding high-quality solutions and finding and proving optimal solutions over a longer run-time.
\end{enumerate}

Our experiments are focused on the performance of the two approaches
compared to 
existing optimization methods. To this end,
we experiment with 12 standard WFLO benchmark problems (c.f.~\cite{turner2014new})
and show that solving quadratic WFLO models
using Gurobi and the DA
achieves state-of-the-art performance.
%, often outperforming existing solutions when provided sufficient run-time.
Furthermore, we observe that the DA often provides solutions of
similar quality to software-based approaches but with two orders of
magnitude shorter run-time.
%The ability to quickly obtain competitive
%solutions on large WFLO instances can be employed to solve multiple
%wind farm design problems.

@@

%\paragraph{Wind Farm as a Grid}
We consider a square wind farm that we model as an $n\times n$
grid with $n$ being the number of cells on each axis.  The cells have
an area of $c^2$ (in square meters), i.e., if $c=100$ meters and
$n=20$, the farm area will be $2km\times 2km$.  We
assume that the terrain is flat and that we have no constraints
related to the terrain.  Furthermore, we do not account for the
influence of noise on the wind farm's surroundings. The wind farm
literature treats the more complex farm models that capture non-flat
terrain~\cite{song2015lazy,kuo2016wind}, multi-typed
turbines~\cite{feng2017design}, noise
effects~\cite{Zhang2014,sorkhabi2016impact,yin2014multi}, and complex
objectives (including turbine installation and maintenance
costs)~\cite{lackner2007analytical,sorkhabi2018constrained}.  In this work, we 
choose to focus on a
novel approach for solving WFLO, and hence consider the most
abstract wind farm model. Extension of our framework to include such additional problem characteristics is left as future research.
%Our model can be extended, in the
%future, to include additional problem characteristics as mentioned above.

@@

\paragraph{Proximity Constraints}
Turbines must be placed at least \emph{five rotor diameters}
apart. Depending on the cell size, we formulate the
corresponding constraint as part of the optimization model. For
example, if a turbine is placed in the center of a
cell, a cell size of $c = 100$ meters and a rotor diameter
of $40$ meters, requires that two turbines are not placed in adjacent cells. In Figure~\ref{fig:field_model}, the red area
around the turbine represents the forbidden cells. Conversely, if $c = 200$ and turbines are placed in the center of the cell, no proximity constraints are required as the required separation is enforced by the granularity of the spatial discretization.

@@

Let $u_{id}$ be the wind velocity at turbine $i = 1,\ldots, m$ for wind regime $d\in\mathcal{D}$.
%\footnote{Turbines can adjust their orientation according to the current wind regime.}
  Given the interference of the wakes, the velocity of turbine $i$ is not equivalent to the free wind speed of $d$.
Rather, $u_{id}$ can be computed using the following equation~\cite{Zhang2014}:
\begin{equation}
u_{id} = u_{id\infty} \Bigg[1 - \sqrt{\sum_{j\in\mathcal{U}_{id}} \bigg( 1-\frac{u_{ijd}}{u_{id\infty}} \bigg)^2}  \Bigg], \label{eq:ss}
\end{equation} with $u_{id\infty}$ being the free wind speed (the second component of regime $d$), $\mathcal{U}_{id}$ corresponding to the set of turbines  that are upstream to $i$ 
under wind regime $d$, and $u_{ijd}$ being the wake-reduced speed at $i$ due to upstream turbine $j$. 
The reduced speed, $u_{ijd}$, is computed using the distance $x$ between the two turbines ($i,j$)  and turbine and terrain specifications that yield the cone opening radius, $r$. The set of upstream turbines, $\mathcal{U}_{id}$ is computed  
by rotating the grid wrt to the current wind regime $d \in \mathcal{D}$ and identifying the turbines, $j$, located upstream of turbine $i$ given $d$. The expression in Eq. (\ref{eq:ss}) is referred to as the \emph{sum-of-squares} (SS) model for 
the total expected power in presence of wakes~\cite{Zhang2014}.  We can now write 
the sum-of-squares (SS) expected power of a wind farm with $m$ turbines:
%~(\cite{Zhang2014}):
\begin{equation}
  E_{SS} = \sum_{i=1}^m \sum_{d\in\mathcal{D}} \frac{1}{3} u_{id}^3p_d.\label{eq:ess}
\end{equation}

@@

To overcome this limitation, Rivas et al.\ \cite{rivas2009solving} used simulated annealing (SA), 
a neighborhood search method that allows visits to infeasible solutions. 
SA creates a well-connected solution space that allows the algorithm to escape 
local optima. The SA method was shown to be superior to the approach of Ozturk \& Norman~\cite{ozturk2004heuristic} 
and to the genetic method proposed in Grady et al.\ \cite{grady2005placement}. A drawback of the SA approach 
is its ad-hoc nature. The various components of the algorithm
 must be tailored to the problem at hand and, as a result, small adjustments of the WFLO setting
 (e.g., adding noise considerations), would require major changes to the SA implementation. 
Additional methods and experimental comparisons between various approximate algorithms are available in Samorani~\cite{samorani2013wind}.
